---
title: "Class Example: Data Import"
output: html_document
---


```{r setup, message = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

## Read in some data

Read in the litters dataset
(fyi:from study on fetal alcohol syndrome with mice)

``` {r}
litters_df = read_csv("./data_import_examples/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
  # In the janitor package, load clean_names function. Since this is the only function we'll use from janitor package, don't bother loading the whole thing
```

## Take a look at the data

``` {r}
litters_df # shows first 10 rows bc we have tidyverse on (or bc we read it in through read_csv)
```

``` {r}
head(litters_df) # shows first 6 rows
```

``` {r}
tail(litters_df) # shows the end of the df
```

``` {r}
skimr::skim(litters_df) # gives a little summary, "gimmiky"
```
 
 You can also use: view(litters_df), which will open up another window. Generally adviseable to put this in the console, bc putting it in a code chunk will make R mkd run a little weird.
 
## Options to read_csv

?read.csv for more info

``` {r, eval=FALSE}
litters_df = read_csv("./data_import_examples/FAS.litters.csv", skip = 10, col_names = FALSE)
# skips first ten rows, not sure what col_names does exactly (doesn't give the columns names, I guess)
litters_df = read_csv("./data_import_examples/FAS.litters.csv", na - c("", "NA", "99"))
# lets read csv know that these are missing values
```

## Other file formats

Read in an excel file

```{r}
mlb_df = read_excel("./data_import_examples/mlb11.xlsx")
```

Options:

"sheet" = pretty excel specific
"range" = gives you options of what range to read in. eg: range = "A1:F7
Can find more options in ?read_excel. Mose are similar to read.csv


Read in a SAS file

```{r}
pulse_df = read_sas("./data_import_examples/public_pulse_data.sas7bdat")
```

## Comparison with Base R

what about read.csv ? He strongly encourages read_csv and NOT read.csv

```{r}
litters_base = read.csv("./data_import_examples/FAS_litters.csv")
litters_readr = read_csv("./data_import_examples/FAS_litters.csv")

litters_base
litters_readr
```

## Exporting data

export the mlb sub-table

```{r, eval=FALSE}
#write_csv(mlb.df, "./data_import_examples/mlb_subtable.csv")
```

Didn't actually eval this because I don't think I actually subsetted the mlb file above the way he did in the example. But this is how this would be done.


